---
title: 'ggplot2 Elegant Graphics: Chapter 10'
author: "Brandon Foltz"
date: "2023-03-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
```{r}
ggplot(mpg, aes(x = displ)) +
  geom_histogram()

# The same as...

ggplot(mpg, aes(x = displ, y = after_stat(count))) +
  geom_histogram()
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(vars(year))
```
```{r}
#different scales issue
mpg_99 <- mpg |>
  filter(year == 1999)
mpg_08 <- mpg|>
  filter(year == 2008)

base_99 <- ggplot(mpg_99, aes(displ, hwy)) + geom_point()
base_08 <- ggplot(mpg_08, aes(displ, hwy)) + geom_point()

base_99
base_08
```
```{r}
base_99 +
  scale_x_continuous(limits = c(1,7)) +
  scale_y_continuous(limits = c(10, 45))

base_08 +
  scale_x_continuous(limits = c(1,7)) +
  scale_y_continuous(limits = c(10, 45))
```

```{r}
base_99 + lims(x = c(1, 7), y = c(10, 45))
base_08 + lims(x = c(1, 7), y = c(10, 45))
```

```{r}
base <- ggplot(mpg, aes(drv, hwy)) +
  geom_hline(yintercept = 28, color = "red") +
  geom_boxplot()

base
base + coord_cartesian(ylim = c(10, 35))
base + ylim(10, 35)
```

```{r}
base <- ggplot(faithfuld, aes(waiting, eruptions)) +
  geom_raster(aes(fill = density)) +
  theme(legend.position = "none") +
  labs(X = NULL, y = NULL)

base
base + 
  scale_x_continuous(expand = expansion(0)) +
  scale_y_continuous(expand = expansion(0))
```

```{r}
base +
  scale_x_continuous(expand = expansion(add = 3)) +
  scale_y_continuous(expand = expansion(add = 3))

# Multiplicative expansion 20%

base +
  scale_x_continuous(expand = expansion(mult = 0.2)) +
  scale_y_continuous(expand = expansion(mult = 0.2))

# Multiplicative expansion of 5% at the lower end of each axes,
# and 20% at the upper end; for the y-axis the expansion is 
# set directly instead of using expansion()

base +
  scale_x_continuous(expand = expansion(mul = c(0.05, 0.2))) +
  scale_y_continuous(expand = c(0.05, 0, 0.2, 0))
```

Breaks

```{r}
toy <- data.frame(
  const = 1,
  up = 1:4,
  txt = letters[1:4],
  big = (1:4)*1000,
  log = c(2, 5, 10, 2000)
)

toy

base <- ggplot(toy, aes(big, const)) +
  geom_point() +
  labs(x = NULL, y = NULL) +
  scale_y_continuous(breaks = NULL)

base
```
```{r}
base + scale_x_continuous(breaks = c(1000, 2000, 4000))
base + scale_x_continuous(breaks = c(1000, 1500, 2000, 4000))
```
You can write your own break function, but in many cases there is no need, thanks to the scales package.32 It provides several tools that are useful for this purpose:

scales::breaks_extended() creates automatic breaks for numeric axes.
scales::breaks_log() creates breaks appropriate for log axes.
scales::breaks_pretty() creates “pretty” breaks for date/times.
scales::breaks_width() creates equally spaced breaks.

```{r}
base
base + scale_x_continuous(breaks = scales::breaks_extended())
base + scale_x_continuous(breaks = scales::breaks_extended(n = 2))
```

```{r}
base +
  scale_x_continuous(breaks = scales::breaks_width(800))
base +
  scale_x_continuous(breaks = scales::breaks_width(800, offset = 200))
base +
  scale_x_continuous(breaks = scales::breaks_width(800, offset = -200))
```

Minor breaks

```{r}
mb <- unique(as.numeric(1:10 %o% 10 ^ (0:3)))
mb

base <- ggplot(toy, aes(log, const)) +
  geom_point() +
  labs(x = NULL, y = NULL) +
  scale_y_continuous(breaks = NULL)

base + scale_x_log10()
base + scale_x_log10(minor_breaks = mb)
```

Labels

```{r}
base <- ggplot(toy, aes(big, const)) +
  geom_point() +
  labs(x = NULL, y = NULL) +
  scale_y_continuous(breaks = NULL)

base
base +
  scale_x_continuous(
    breaks = c(2000, 4000),
    labels = c("2k", "4k")
  )
```
Again, the scales package provides a number of tools that will automatically construct label functions for you. Some of the more useful examples for numeric data include:

scales::label_bytes() formats numbers as kilobytes, megabytes etc.
scales::label_comma() formats numbers as decimals with commas added.
scales::label_dollar() formats numbers as currency.
scales::label_ordinal() formats numbers in rank order: 1st, 2nd, 3rd etc.
scales::label_percent() formats numbers as percentages.
scales::label_pvalue() formats numbers as p-values: <.05, <.01, .34, etc.

```{r}
base <- ggplot(toy, aes(big, const)) +
  geom_point() +
  labs(x = NULL, y = NULL) +
  scale_x_continuous(breaks = NULL)

base
base + scale_y_continuous(labels = scales::label_percent())
base + scale_y_continuous(
  labels = scales::label_dollar(prefix = "$", suffix = "")
)
```
Transformations

```{r}
base <- ggplot(mpg, aes(displ, hwy)) + geom_point()

base
base + scale_x_reverse()
base + scale_y_reverse()
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(trans = "reciprocal")

ggplot(diamonds, aes(price, carat)) +
  geom_bin_2d() +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```

```{r}
#equivalent

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(trans = "reciprocal")

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(trans = scales::reciprocal_trans())
```
```{r}
#equivalent

ggplot(diamonds, aes(price, carat)) + 
  geom_bin2d() + 
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")

ggplot(diamonds, aes(price, carat)) + 
  geom_bin2d() + 
  scale_x_log10() +
  scale_y_log10()
```

```{r}
# the axis is different for each
ggplot(mpg, aes(log10(displ), hwy)) +
  geom_point()

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_log10()
```
Date-time Position Scales

```{r}
date_base <- ggplot(economics, aes(date, psavert)) +
  geom_line(na.rm = TRUE) +
  labs(x = NULL, y = NULL)

date_base
date_base + scale_x_date(date_breaks = "15 years")
```

```{r}
the_year <- as.Date(c("2021-01-01", "2021-12-31"))
set_breaks <- scales::breaks_width("1 month")
set_breaks(the_year)
```

```{r}
set_breaks <- scales::breaks_width("1 month", offset = 8)
set_breaks(the_year)
```

```{r}
df <- data.frame(
  y = as.Date(c("2022-01-01", "2022-04-01")))
base <- ggplot(df, aes(y = y)) +
  labs(y = NULL) +
  theme_minimal() +
  theme(
    panel.grid.major = element_line(color = "black"),
    panel.grid.minor = element_line(color = "grey50")
  )

base + scale_y_date(date_breaks = "1 month")
base + scale_y_date(date_breaks = "1 month", date_minor_breaks = "1 week")
```

String	Meaning
%S	second (00-59)
%M	minute (00-59)
%l	hour, in 12-hour clock (1-12)
%I	hour, in 12-hour clock (01-12)
%p	am/pm
%H	hour, in 24-hour clock (00-23)
%a	day of week, abbreviated (Mon-Sun)
%A	day of week, full (Monday-Sunday)
%e	day of month (1-31)
%d	day of month (01-31)
%m	month, numeric (01-12)
%b	month, abbreviated (Jan-Dec)
%B	month, full (January-December)
%y	year, without century (00-99)
%Y	year, with century (0000-9999)

```{r}
base <- ggplot(economics, aes(date, psavert)) +
  geom_line(na.rm = TRUE) +
  labs(x = NULL, y = NULL)

base + scale_x_date(date_breaks = "5 years")
base + scale_x_date(date_breaks = "5 years", date_labels = "%y")
```
```{r}
# line breaks
lim <- as.Date(c("2004-01-01", "2005-01-01"))

base + scale_x_date(limits = lim, date_labels = "%b %y")
base + scale_x_date(limits = lim, date_labels = "%B\n%Y")
```

```{r}
base +
  scale_x_date(
    limits = lim,
    labels = scales::label_date_short()
  )
```
Discrete Position Scales

```{r}
ggplot(mpg, aes(x = hwy, y = class)) +
  geom_point()

ggplot(mpg, aes(x = hwy, y = class)) +
  geom_point() +
  scale_x_continuous() +
  scale_y_discrete()
```
```{r}
ggplot(mpg, aes(x = hwy, y = class)) +
  geom_point() +
  annotate("text", x = 5, y = 1:7, label = 1:7)

ggplot(mpg, aes(x = hwy, y = class)) +
  geom_jitter(width = 0, height = 0.25) +
  annotate("text", x = 5, y = 1:7, label = 1:7)
```

```{r}
ggplot(mpg, aes(x = drv, y = hwy)) + geom_boxplot()
ggplot(mpg, aes(x = drv, y = hwy)) + geom_boxplot(width = 0.4)
```

```{r}
base <- ggplot(toy, aes(const, txt)) +
  geom_label(aes(label = txt)) +
  scale_x_continuous(breaks = NULL) +
  labs(x = NULL, y = NULL)

base
base + scale_y_discrete(limits = c("a", "b", "c", "d", "e"))
base + scale_y_discrete(limits = c("d", "c", "a", "b"))
```
```{r}
base + scale_y_discrete(breaks = c("b", "c"))
base + scale_y_discrete(labels = c(c = "carrot", b = "banana")) 
```
Label Positions

```{r}
base <- ggplot(mpg, aes(manufacturer, hwy)) +
  geom_boxplot()
base

base + guides(x = guide_axis(angle = 45))
```

Binned Position Scales

```{r}
ggplot(mpg, aes(hwy)) +
  geom_histogram(bins = 8)
ggplot(mpg, aes(hwy)) +
  geom_bar() +
  scale_x_binned()
```

```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_count()

base

base +
  scale_x_binned(n.breaks = 15) +
  scale_y_binned(n.breaks = 15)
```

