---
title: 'ggplot2 Elegant Graphics: Chapter 11'
author: "Brandon Foltz"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
colorBlindness::displayAllColors(rainbow(6))
colorBlindness::displayAllColors(viridis::viridis(6))
```

```{r}
erupt <- ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_raster() +
  scale_x_continuous(NULL, expand = c(0,0)) +
  scale_y_continuous(NULL, expand = c(0,0)) +
  theme(legend.position = "none")

erupt
erupt + scale_fill_viridis_c()
erupt + scale_fill_viridis_c(option = "magma")
```

```{r}
erupt
erupt + scale_fill_distiller()
erupt + scale_fill_distiller(palette = "RdPu")
erupt + scale_fill_distiller(palette = "YlOrBr")
```

```{r}
erupt + scico::scale_fill_scico(palette = "bilbao") # the default
erupt + scico::scale_fill_scico(palette = "vik")
erupt + scico::scale_fill_scico(palette = "lajolla")
```

```{r}
erupt + paletteer::scale_fill_paletteer_c("viridis::plasma")
erupt + paletteer::scale_fill_paletteer_c("scico::tokyo")
```

scale_fill_gradient() produces a two-colour gradient
scale_fill_gradient2() produces a three-colour gradient with specified midpoint
scale_fill_gradientn() produces an n-colour gradient

```{r}
erupt + scale_fill_gradient(low = "grey", high = "brown")
erupt + 
  scale_fill_gradient2(
    low = "grey", 
    mid = "white", 
    high = "brown", 
    midpoint = .02
  )
erupt + scale_fill_gradientn(colours = terrain.colors(7))
```

```{r}
munsell::hue_slice("5P") +  # Generate a ggplot with hue_slice()
  annotate(                 # Add arrows for annotation 
    geom = "segment", 
    x = c(7, 7), 
    y = c(1, 10), 
    xend = c(7, 7), 
    yend = c(2, 9), 
    arrow = arrow(length = unit(2, "mm"))
  ) 
#> Warning: Removed 31 rows containing missing values (`geom_text()`).

# Construct scale
erupt + scale_fill_gradient(
  low = munsell::mnsl("5P 2/12"), 
  high = munsell::mnsl("5P 7/12")
)
```

```{r}
# munsell example
erupt + scale_fill_gradient2(
  low = munsell::mnsl("5B 7/8"),
  high = munsell::mnsl("5Y 7/8"),
  mid = munsell::mnsl("N 7/0"),
  midpoint = .02
) 

# colorspace examples
erupt + scale_fill_gradientn(colours = colorspace::heat_hcl(7))
erupt + scale_fill_gradientn(colours = colorspace::diverge_hcl(7))
```

```{r}
df <- data.frame(x = 1, y = 1:5, z = c(1, 3, 2, NA, 5))
base <- ggplot(df, aes(x, y)) + 
  geom_tile(aes(fill = z), size = 5) + 
  labs(x = NULL, y = NULL) +
  scale_x_continuous(labels = NULL)

base
base + scale_fill_gradient(na.value = NA)
base + scale_fill_gradient(na.value = "yellow")
```

```{r}
# base <- ggplot(toy, aes(up, up, fill = big)) + 
#   geom_tile() + 
#   labs(x = NULL, y = NULL) 
# 
# base 
# base + scale_fill_continuous(limits = c(0, 10000))
```

```{r}
base <- ggplot(mpg, aes(cyl, displ, colour = hwy)) +
  geom_point(size = 2)

base

base + guides(colour = guide_colourbar(reverse = TRUE))
base + guides(colour = guide_colourbar(barheight = unit(2, "cm")))
base + guides(colour = guide_colourbar(direction = "horizontal"))
```

```{r}
df <- data.frame(x = c("a", "b", "c", "d"), y = c(5, 1.75, 3, .5))

bars <- ggplot(df,
       aes(x, y, fill = x)) +
  geom_bar(stat = "identity") +
  labs(x=NULL, y = NULL) +
  theme(legend.position = "none")

bars
```

```{r}
# RColorBrewer::display.brewer.all()

bars + scale_fill_brewer(palette = "Set1")
bars + scale_fill_brewer(palette = "Set2")
bars + scale_fill_brewer(palette = "Accent")
```

```{r}
# scatter plot
df <- data.frame(
  x = 1:3 + runif(90), 
  y = runif(90), 
  z = c("a", "b", "c")
)
point <- ggplot(df, aes(x, y)) +
  geom_point(aes(colour = z))  + 
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)

# three palettes
point + scale_colour_brewer(palette = "Set1")
point + scale_colour_brewer(palette = "Set2")  
point + scale_colour_brewer(palette = "Pastel1")
```

```{r}
# bar plot
df <- data.frame(x = 1:3, y = 3:1, z = c("a", "b", "c"))
area <- ggplot(df, aes(x, y)) + 
  geom_bar(aes(fill = z), stat = "identity") + 
  theme(legend.position = "none") +
  labs(x = NULL, y = NULL)

# three palettes
area + scale_fill_brewer(palette = "Set1")
area + scale_fill_brewer(palette = "Set2")
area + scale_fill_brewer(palette = "Pastel1")
```

You can control the default chroma and luminance, and the range of hues, with the h, c and l arguments:
```{r}
bars
bars + scale_fill_hue(c = 40)
bars + scale_fill_hue(h = c(180, 300))
```

```{r}
bars + scale_fill_grey()
bars + scale_fill_grey(start = 0.5, end = 1)
bars + scale_fill_grey(start = 0, end = 0.5)
```

```{r}
bars + paletteer::scale_fill_paletteer_d("rtist::vangogh")
bars + paletteer::scale_fill_paletteer_d("colorBlindness::paletteMartin")
bars + paletteer::scale_fill_paletteer_d("wesanderson::FantasticFox1")

```

Manual scales

```{r}
bars + 
  scale_fill_manual(
    values = c("sienna1", "sienna4", "hotpink1", "hotpink4")
  )

bars + 
  scale_fill_manual(
    values = c("tomato1", "tomato2", "tomato3", "tomato4")
  )

bars + 
  scale_fill_manual(
    values = c("grey", "black", "grey", "grey")
  )
```

```{r}
bars + 
  scale_fill_manual(
    values = c(
      "d" = "grey",
      "c" = "grey",
      "b" = "black",
      "a" = "grey"
    )
  )
```

```{r}
mpg_99 <- mpg %>% filter(year == 1999)
mpg_08 <- mpg %>% filter(year == 2008)

base_99 <- ggplot(mpg_99, aes(displ, hwy, colour = fl)) + geom_point() 
base_08 <- ggplot(mpg_08, aes(displ, hwy, colour = fl)) + geom_point() 

base_99
base_08
```

```{r}
base_99 + lims(colour = c("c", "d", "e", "p", "r"))
base_08 + lims(colour = c("c", "d", "e", "p", "r"))
```

```{r}
base_99 + 
  lims(
    x = c(1, 7), 
    y = c(10, 45), 
    colour = c("c", "d", "e", "p", "r")
  )

base_08 + 
  lims(
    x = c(1, 7), 
    y = c(10, 45), 
    colour = c("c", "d", "e", "p", "r")
  )
```

Legends

```{r}
base <- ggplot(mpg, aes(drv, fill = factor(cyl))) + geom_bar() 

base
base + guides(fill = guide_legend(ncol = 2))
base + guides(fill = guide_legend(ncol = 2, byrow = TRUE))
```

```{r}
base
base + guides(fill = guide_legend(reverse = TRUE))
```

Override 
```{r}
base <- ggplot(mpg, aes(displ, hwy, colour = drv)) +
  geom_point(size = 4, alpha = .2, stroke = 0)

base + guides(colour = guide_legend())
base + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

Binned Color Scale

```{r}
erupt + scale_fill_binned()
erupt + scale_fill_steps()
erupt + scale_fill_steps(n.breaks = 8)
```

```{r}
erupt + scale_fill_steps(low = "grey", high = "brown")
erupt + 
  scale_fill_steps2(
    low = "grey", 
    mid = "white", 
    high = "brown", 
    midpoint = .02
  )
erupt + scale_fill_stepsn(n.breaks = 12, colours = terrain.colors(12))
```

```{r}
erupt + scale_fill_stepsn(n.breaks = 9, colours = viridis::viridis(9))
erupt + scale_fill_stepsn(n.breaks = 9, colours = viridis::magma(9))
erupt + scale_fill_stepsn(n.breaks = 9, colours = viridis::inferno(9))
```

```{r}
erupt + scale_fill_fermenter(n.breaks = 9)
erupt + scale_fill_fermenter(n.breaks = 9, palette = "Oranges")
erupt + scale_fill_fermenter(n.breaks = 9, palette = "PuOr")
```

Legends
```{r}
base <- ggplot(mpg, aes(cyl, displ, colour = hwy)) +
  geom_point(size = 2) +
  scale_color_binned()

base 
base + guides(colour = guide_coloursteps(show.limits = TRUE))
```

Date-time Color

```{r}
base <- ggplot(economics, aes(psavert, uempmed, colour = date)) + 
  geom_point() 

base
base + 
  scale_colour_date(
    date_breaks = "142 months", 
    date_labels = "%b %Y"
  )
```

Alpha Scale

```{r}
ggplot(faithfuld, aes(waiting, eruptions, alpha = density)) +
  geom_raster(fill = "maroon") +
  scale_x_continuous(expand = c(0, 0)) + 
  scale_y_continuous(expand = c(0, 0))

```

Legend Position

```{r}
# base <- ggplot(toy, aes(up, up)) + 
#   geom_point(aes(colour = txt), size = 3) + 
#   xlab(NULL) + 
#   ylab(NULL)
# 
# base + theme(legend.position = "left")
# base + theme(legend.position = "right") # the default 
# base + theme(legend.position = "bottom")
# base + theme(legend.position = "none")

```

